% \documentclass[../main.tex]{subfiles}
% \begin{document}

% \begin{center}
%     \Large{\textbf{ABSTRACT}}\\
% \end{center}
% \vspace{1cm}
% Capacitive sensing technology has become a standard for touch-based interaction, yet most systems treat all finger contacts as uniform inputs. Recognizing and classifying specific finger types (e.g., thumb, index, or middle finger) from capacitive images can significantly enhance human-computer interaction and device ergonomics. However, implementing such recognition models on battery-constrained edge devices requires a strict balance between classification accuracy and power efficiency. This thesis investigates the application of Spiking Neural Networks (SNNs)—a bio-inspired, event-driven computing paradigm—as an energy-efficient solution for capacitive image-based finger recognition.

% The research begins by proposing specialized data exploration methods for the CapfingerId dataset to extract distinctive spatial features of different finger types. To establish a performance benchmark, a standard Convolutional Neural Network (CNN) is developed, followed by the design of an equivalent SNN architecture to allow for a direct comparison of spiking dynamics. Furthermore, the study explores the integration of transformer-based global attention through a Spiking Vision Transformer (Spiking ViT) model. The proposed architectures are rigorously evaluated by measuring their recognition accuracy across various finger categories and their corresponding energy footprints. \textcolor{red}{[Insert your results here: e.g., The results show that the Spiking ViT achieves an accuracy of X\% while consuming only Y mJ per inference]}. This study demonstrates that SNNs provide a viable, low-power alternative to traditional deep learning models for finger-type classification in next-generation interactive devices.
% \begin{flushright}
% \begin{tabular}{@{}c@{}}
% Student\\
% \textit{(Signature and full name)}
% \end{tabular}
% \end{flushright}

% \end{document}

\documentclass[../main.tex]{subfiles}
\begin{document}

\begin{center}
    \Large{\textbf{ABSTRACT}}\\
\end{center}
\vspace{1cm}

Capacitive sensing technology has become a standard for touch-based interaction, yet most systems treat all finger contacts as uniform inputs. Recogning and classifying specific finger types (e.g., thumb, index, or middle finger) from capacitive images can significantly enhance human-computer interaction and device ergonomics. However, deploying such recognition models on battery-constrained edge devices requires a careful balance between classification accuracy and power efficiency.

This thesis investigates the development of efficient deep learning models for capacitive image-based finger recognition, with a particular focus on Spiking Neural Networks (SNNs)—a bio-inspired and event-driven computing paradigm known for its low power consumption and hardware efficiency. Two SNN-based architectures are developed and evaluated in this study.

First, a RANC-based SNN model is proposed as a fully spiking architecture trained entirely from scratch using SNN principles. The model is designed within the Training, Testing, and Emulation environment provided by RANC, enabling strong potential for deployment on neuromorphic and hardware-accelerated platforms such as FPGAs. The proposed RANC-based SNN demonstrates recognition performance comparable to the CNN baseline introduced in the original CapFingerID study, while significantly reducing computational complexity and model parameters.

Second, a Spiking Vision Transformer (Spiking ViT) architecture is explored to incorporate the powerful representation learning capability of Vision Transformers into the spiking domain. The proposed model adopts a hybrid CNN-SNN design, combining efficient spiking computation with global attention mechanisms to improve feature extraction and classification performance. Experimental results show that the Spiking ViT outperforms the original CapFingerID baseline model by approximately $6\%$ in classification accuracy.

Overall, the findings demonstrate that SNN-based architectures—particularly transformer-enhanced spiking models—provide a promising and energy-efficient alternative to conventional deep learning approaches for finger-type recognition in next-generation interactive and embedded systems.
\begin{flushright}
\begin{tabular}{@{}c@{}}
Student\\
\textit{(Signature and full name)}
\end{tabular}
\end{flushright}

\end{document}