\documentclass[../main.tex]{subfiles}
\begin{document}

This chapter presents the experimental evaluation of the proposed RANC-based Spiking Neural Network (SNN) and Spiking Vision Transformer (Spiking ViT) architectures for capacitive fingerprint recognition. The experiments investigate the influence of spatial segmentation, Tea layer output dimensionality, transformer depth, embedding dimension, and temporal encoding windows on both recognition performance and model complexity. For the RANC-based SNN, the results demonstrate that overlapping parallel segmentation and moderate temporal encoding significantly improve classification accuracy while maintaining lightweight computational complexity. In particular, the \((\textit{Elems}=16, \textit{Overlap}=8)\) configuration achieved the most favorable balance between recognition capability and parameter efficiency. The experiments further show that increasing the temporal encoding window improves spike-based feature representation, although the performance gain gradually saturates at larger timestep values. For the Spiking Vision Transformer, deeper encoder structures and larger embedding dimensions consistently improved classification performance through enhanced global feature modeling. Overall, the experimental findings validate the effectiveness of combining spiking computation, temporal encoding, and transformer-based attention mechanisms for efficient and accurate capacitive fingerprint recognition.

\section{Experimental Evaluation of RANC-based SNN}
\label{sec:Experimental-Evaluation-of-RANC-based-SNN}
\subsection{Experimental Setup}

This section presents the experimental configuration used to evaluate the proposed RANC-based Spiking Neural Network (SNN). The experiments were designed to investigate the effects of temporal encoding and architectural parameters on both recognition performance and model complexity.

\subsubsection{Experimental Settings}

The experimental analysis focused on two main groups of parameters:

\begin{itemize}
    \item \textbf{Temporal encoding parameter}: the number of encoding time steps \(T\), which determines the temporal resolution of the burst encoding scheme.
    
    \item \textbf{Parallel segmentation parameters}: including the segment size (\textit{Elems}), overlap size (\textit{Overlap}), and the number of output neurons in each Tea layer (\textit{Tea Out}).
\end{itemize}

To systematically analyze the effect of each parameter, only one parameter group was varied at a time while the remaining parameters were fixed to their default values.

The default configuration of the proposed architecture is summarized in Table~\ref{tab:ranc_default_config}.

\begin{table}[h]
\centering
\caption{Default configuration of the proposed RANC-based SNN}
\label{tab:ranc_default_config}
\begin{tabular}{|c|c|}
\hline
\textbf{Parameter} & \textbf{Value} \\ \hline
Optimizer & Adam \\ \hline
Learning Rate & 0.001 \\ \hline
Maximum Epochs & 300 \\ \hline
Early Stopping Patience & 10 \\ \hline
Dropout Rate & 0.15 \\ \hline
Encoding Window \(T\) & 4 \\ \hline
Segment Size (Elems) & 16 \\ \hline
Overlap & 8 \\ \hline
Tea Out & 256 \\ \hline
\end{tabular}
\end{table}

\subsection{Effect of Parallel Segmentation Strategy}

This experiment investigates the effect of the proposed parallel segmentation strategy on the performance of the RANC-based SNN architecture. In particular, the analysis focuses on two important segmentation parameters: the segment size (\textit{Elems}) and the overlap size (\textit{Overlap}). The experiments were conducted using a temporal encoding window of \(T=1\), corresponding to the binary spike encoding setting.

The objective of this experiment is to evaluate how different segmentation configurations influence feature extraction capability, spatial continuity preservation, and model complexity. Different combinations of segment size and overlap were therefore explored while monitoring both classification performance and the number of trainable parameters.

Table~\ref{tab:ranc-results} summarizes the experimental results obtained for different model configurations.

\begin{table*}[t]
\centering
\caption{Model configurations and performance under different parallel segmentation settings (\(T=1\)).}
\resizebox{0.92\textwidth}{!}{
\begin{tabular}{cccccccc}
\toprule
\multicolumn{3}{c}{\textbf{Model}} & \textbf{Index L/R} & \textbf{Thumb L/R} & \textbf{Thumb/Index} & \textbf{Thumb/Others} & \textbf{5 Fingers}\\
\cmidrule(lr){1-3}
Elems & Overlap & Tea Out & & & & &\\
\midrule
16 & 8 & 64 & 0.60 & 0.84 & 0.78 & 0.81 & 0.42 \\
32 & 8 & 64 & 0.56 & 0.77 & 0.75 & 0.77 & 0.39 \\
16 & 4 & 32 & 0.57 & 0.77 & 0.74 & 0.76 & 0.40 \\
32 & 16 & 128 & 0.61 & 0.84 & 0.79 & 0.82 & 0.43 \\
16 & 4 & 64 & 0.59 & 0.84 & 0.77 & 0.80 & 0.41 \\
32 & 16 & 64 & 0.57 & 0.81 & 0.76 & 0.79 & 0.40 \\
16 & 8 & 128 & 0.62 & 0.85 & \textbf{0.80} & 0.83 & 0.41 \\
32 & 8 & 256 & 0.61 & 0.84 & 0.79 & 0.82 & 0.43 \\
8 & 2 & 64 & 0.57 & 0.81 & 0.75 & 0.78 & 0.40 \\
8 & 2 & 256 & 0.60 & 0.84 & 0.78 & 0.81 & 0.42 \\
32 & 8 & 128 & 0.58 & 0.82 & 0.76 & 0.79 & 0.41 \\
8 & 4 & 64  & 0.59 & 0.83 & 0.77 & 0.80 & 0.41 \\
16 & 4 & 256 & 0.62 & \textbf{0.86} & 0.81 & \textbf{0.84} & 0.42 \\
8 & 4 & 32  & 0.56 & 0.76 & 0.74 & 0.76 & 0.39 \\
32 & 16 & 256 & 0.62 & \textbf{0.86} & \textbf{0.80} & 0.83 & 0.42 \\
16 & 8  & 32  & 0.57 & 0.80 & 0.75 & 0.78 & 0.40 \\
32 & 8  & 32  & 0.55 & 0.62 & 0.74 & 0.76 & 0.39 \\
8  & 2  & 32  & 0.56 & 0.79 & 0.75 & 0.77 & 0.40 \\
32 & 16 & 32  & 0.55 & 0.66 & 0.74 & 0.76 & 0.39 \\
8  & 4  & 256 & 0.61 & 0.85 & \textbf{0.80} & \textbf{0.84} & 0.43 \\
8  & 2  & 128 & 0.58 & 0.82 & 0.76 & 0.79 & 0.41 \\
16 & 4  & 128 & 0.61 & 0.85 & 0.79 & 0.82 & 0.43 \\
\textbf{16} & \textbf{8}  & \textbf{256} & \textbf{0.63} & \textbf{0.86} & 0.79 & 0.82 & \textbf{0.44} \\
8  & 4  & 128 & 0.60 & 0.85 & 0.78 & 0.81 & 0.42 \\
\bottomrule  
\end{tabular}
}
\label{tab:ranc-results}
\end{table*}

\begin{table*}[t]
\centering
\caption{Number of trainable parameters under different parallel segmentation configurations.}
\begin{tabular}{cccc}
\toprule
\multicolumn{3}{c}{\textbf{Model}} & \textbf{\#Params} \\
\cmidrule(lr){1-3}
Elems & Overlap & Tea Out &\\
\midrule
16 & 8 & 64 &14784 \\
32 & 8 & 64 &8320 \\
16 & 4 & 32 &5280 \\
32 & 16 & 128 &24960 \\
16 & 4 & 64 &10560 \\
32 & 16 & 64 &12480 \\
16 & 8 & 128 &29568 \\
32 & 8 & 256 &33280 \\
8 & 2 & 64 &10880 \\
8 & 2 & 256 &43520 \\
32 & 8 & 128 &16640 \\
8 & 4 & 64  &16320 \\
16 & 4 & 256 &42240 \\
8 & 4 & 32  &8160  \\
32 & 16 & 256 &49920 \\
16 & 8  & 32  &7392  \\
32 & 8  & 32  &4160  \\
8  & 2  & 32  &5440  \\
32 & 16 & 32  &6240  \\
8  & 4  & 256 &65280 \\
8  & 2  & 128 &21760 \\
16 & 4  & 128 &21120 \\
16 & 8  & 256 &59136 \\
8  & 4  & 128 &32640 \\
\bottomrule
\end{tabular}
\label{tab:num-params}
\end{table*}

The experimental results demonstrate that the proposed parallel segmentation strategy significantly affects the recognition capability of the RANC-based SNN architecture across all fingerprint classification tasks. In general, configurations with moderate segment sizes and sufficient overlap consistently achieved better classification accuracy compared to configurations using either excessively large segments with limited overlap or very small fragmented segments.

Among all evaluated configurations, the $(16,8)$ segmentation setting consistently produced the strongest overall performance. Under \textit{Tea Out}=256, this configuration achieved the best results in several tasks, including $0.63$ for \textit{Index L/R}, $0.86$ for \textit{Thumb L/R}, and the highest $5$-finger classification accuracy of $0.44$, while using $59136$ trainable parameters. Even with smaller output dimensions, the $(16,8)$ setting remained highly stable, achieving $0.60$, $0.62$, and $0.63$ for \textit{Index L/R} under \textit{Tea Out}=64, 128, and 256, respectively. These results indicate that the $(16,8)$ configuration provides an effective balance between localized feature extraction and preservation of spatial continuity across neighboring fingerprint regions.

In contrast, configurations using larger segment sizes with insufficient overlap, particularly $(32,8)$, consistently produced weaker performance. For example, the $(32,8,32)$ configuration achieved only $0.55$ for \textit{Index L/R}, $0.62$ for \textit{Thumb L/R}, $0.74$ for \textit{Thumb/Index}, and $0.39$ for the $5$-finger classification task, despite using only $4160$ trainable parameters. Although increasing \textit{Tea Out} improved performance for this segmentation setting, the overall accuracy remained lower than configurations with larger overlap. For instance, $(32,8,128)$ achieved $0.58$ for \textit{Index L/R} and $0.82$ for \textit{Thumb L/R}, which were still below the corresponding results obtained by the $(32,16)$ configuration.

The importance of overlap can be observed clearly when comparing $(32,8)$ and $(32,16)$. Increasing the overlap from 8 to 16 improved \textit{Thumb L/R} accuracy from $0.77$ to $0.81$ for \textit{Tea Out}=64, and from $0.82$ to $0.84$ for \textit{Tea Out}=128. Similarly, \textit{Index L/R} improved from $0.56$ to $0.57$ for \textit{Tea Out}=64 and from $0.58$ to $0.61$ for \textit{Tea Out}=128. The $5$-finger classification accuracy also increased from $0.39$ to $0.40$ and $0.41$ to $0.43$, respectively. These improvements suggest that overlapping neighboring regions helps preserve shared spatial patterns and reduces information loss near segmentation boundaries.

Configurations using very small segments with limited overlap, such as $(8,2)$, also exhibited reduced performance in several cases. Under \textit{Tea Out}=32, the model achieved only $0.56$ for \textit{Index L/R}, $0.79$ for \textit{Thumb L/R}, and $0.40$ for the $5$-finger task, despite requiring $5440$ parameters. Although increasing the output dimension to 256 improved the performance to $0.60$, $0.84$, and $0.42$, the results remained slightly lower than those obtained by the $(16,8)$ configuration. This behavior suggests that excessive fragmentation of the input reduces the contextual information available to each processing branch, limiting the model’s ability to learn discriminative fingerprint representations.

A similar overlap-related improvement can also be observed for smaller segmentation settings. Comparing $(8,2)$ and $(8,4)$, increasing the overlap improved \textit{Thumb L/R} accuracy from $0.81$ to $0.83$ for \textit{Tea Out}=64, and from $0.82$ to $0.85$ for \textit{Tea Out}=128. The \textit{Thumb/Others} classification accuracy also improved from $0.78$ to $0.80$ for \textit{Tea Out}=64 and from $0.79$ to $0.81$ for \textit{Tea Out}=128. These observations further confirm that overlapping segmentation contributes positively to feature continuity and classification robustness.

From a computational perspective, the segmentation strategy also strongly affects model complexity. Increasing overlap generally introduces additional trainable parameters because of the larger effective receptive regions processed by each Tea layer. For example, under \textit{Tea Out}=128, increasing overlap from 8 to 16 for the 32-element segmentation setting increased the parameter count from $16640$ to $24960$. Similarly, for the 8-element segmentation setting, increasing overlap from 2 to 4 increased the number of parameters from $21760$ to $32640$. Moreover, increasing \textit{Tea Out} substantially enlarged the model size, as observed in the $(16,8)$ configuration where the parameter count increased from $7392$ at \textit{Tea Out}=32 to $59136$ at \textit{Tea Out}=256.

Overall, the experimental results demonstrate that the proposed overlapping parallel segmentation strategy effectively improves feature extraction capability and classification stability across multiple fingerprint classification tasks. Among all evaluated configurations, moderate segment sizes with moderate overlap, particularly the $(16,8)$ configuration, provide the most favorable balance between recognition performance and computational complexity.

\subsection{Effect of Tea Out}
This experiment investigates the influence of the \textit{Tea Out} parameter on the representational capability and classification performance of the proposed parallel RANC-based SNN architecture. The \textit{Tea Out} parameter determines the dimensionality of the output feature representation generated by each Tea layer. Increasing this value enlarges the latent spiking feature space and allows the network to encode more discriminative fingerprint characteristics.

The experiments were conducted under multiple segmentation configurations while varying the \textit{Tea Out} dimension among \(32\), \(64\), \(128\), and \(256\). The objective is to analyze the trade-off between classification performance and model complexity as the output dimensionality increases.

The experimental results in Table~\ref{tab:ranc-results} show that increasing the \textit{Tea Out} dimension generally improves recognition performance across all evaluation tasks. In most configurations, larger output dimensions consistently achieved higher scores for \textit{Index L/R}, \textit{Thumb L/R}, \textit{Thumb/Index}, \textit{Thumb/Others}, and \textit{5 Fingers} classification tasks.

For the \((16,8)\) segmentation configuration, increasing \textit{Tea Out} from \(32\) to \(64\), \(128\), and \(256\) improved the \textit{Index L/R} score from \(0.57\) to \(0.60\), \(0.62\), and \(0.63\), respectively. Similarly, the \textit{Thumb L/R} performance increased from \(0.80\) to \(0.84\), \(0.85\), and finally \(0.86\). The \textit{5 Fingers} evaluation also improved from \(0.40\) to \(0.42\), \(0.41\), and \(0.44\). These results indicate that larger Tea representations enable the model to capture richer local fingerprint structures and improve classification robustness.

A similar trend can also be observed for the \((32,16)\) configuration. When \textit{Tea Out}=32, the model achieved relatively low scores, including \(0.55\) for \textit{Index L/R} and \(0.66\) for \textit{Thumb L/R}. Increasing the output dimension to \(64\), \(128\), and \(256\) progressively improved the performance to \(0.57\), \(0.61\), and \(0.62\) for \textit{Index L/R}, while \textit{Thumb L/R} improved to \(0.81\), \(0.84\), and \(0.86\), respectively. The \textit{Thumb/Others} metric also improved from \(0.76\) to \(0.79\), \(0.82\), and \(0.83\).

For smaller segmentation settings such as \((8,2)\), the same behavior was consistently observed. Increasing \textit{Tea Out} from \(32\) to \(64\), \(128\), and \(256\) improved the \textit{Thumb L/R} score from \(0.79\) to \(0.81\), \(0.82\), and \(0.84\), respectively. Likewise, the \textit{Thumb/Index} metric improved from \(0.75\) to \(0.75\), \(0.76\), and \(0.78\). Although the improvement becomes smaller at larger output dimensions, higher-dimensional Tea representations still provide more stable classification performance across multiple evaluation tasks.

The results suggest that increasing the output dimensionality enables each Tea branch to encode richer localized spiking features. Since the proposed architecture relies on parallel feature extraction from segmented fingerprint regions, larger feature representations provide additional flexibility for capturing subtle ridge patterns and spatial variations in capacitive fingerprint images.

However, increasing the \textit{Tea Out} dimension also substantially increases the number of trainable parameters, as shown in Table~\ref{tab:num-params}. For the \((16,8)\) configuration, the parameter count increased from \(7392\) at \textit{Tea Out}=32 to \(14784\), \(29568\), and \(59136\) for output dimensions of \(64\), \(128\), and \(256\), respectively. Similarly, under the \((8,4)\) configuration, the number of parameters increased from \(8160\) to \(16320\), \(32640\), and \(65280\).

Although larger output dimensions generally improve performance, the improvement becomes progressively smaller beyond \textit{Tea Out}=128. For example, under the \((16,8)\) configuration, increasing \textit{Tea Out} from \(128\) to \(256\) improved the \textit{Index L/R} score by only \(0.01\) (\(0.62 \rightarrow 0.63\)) and the \textit{Thumb L/R} score by only \(0.01\) (\(0.85 \rightarrow 0.86\)), while the number of trainable parameters doubled from \(29568\) to \(59136\). Similar diminishing returns can also be observed in the \((32,16)\) and \((8,2)\) configurations.

Overall, the experimental results demonstrate that the \textit{Tea Out} parameter plays a critical role in balancing representational capability and computational complexity. Small output dimensions often lead to insufficient feature representation and reduced classification performance, while excessively large output dimensions significantly increase model complexity with relatively limited performance gain. Among the evaluated settings, moderate-to-large output dimensions such as \(128\) and \(256\) provide the most favorable balance between recognition accuracy and computational efficiency for the proposed parallel RANC-based SNN architecture.

\subsection{Effect of Temporal Encoding Window}

This experiment investigates the effect of the temporal encoding window \(T\) on the performance of the proposed RANC-based SNN architecture. To isolate the influence of temporal encoding, the experiments were conducted using the best-performing segmentation configuration identified in the previous experiments, namely \textit{Elems}=16, \textit{Overlap}=8, and \textit{Tea Out}=256.

The temporal encoding window determines the number of discrete time steps used in the burst encoding process. Increasing \(T\) allows each pixel intensity to be represented with finer temporal resolution through longer spike bursts, potentially enabling the network to capture richer spatio-temporal information.

Table~\ref{tab:timestep_results} summarizes the experimental results obtained under different temporal encoding windows.

\begin{table}[h]
\centering
\caption{Performance of the proposed RANC-based SNN under different temporal encoding windows using the best segmentation configuration (\textit{Elems}=16, \textit{Overlap}=8, \textit{Tea Out}=256).}
\resizebox{0.92\textwidth}{!}{
\label{tab:timestep_results}
\begin{tabular}{cccccc}
\toprule
\textbf{Time Steps (\(T\))} & \textbf{Index L/R} & \textbf{Thumb L/R} & \textbf{Thumb/Index} & \textbf{Thumb/Others} & \textbf{5 Fingers} \\
\midrule
1  & 0.63 & 0.86 & 0.79 & 0.82 & 0.44\\
2  & 0.63 & 0.87 & 0.80 & 0.83 & 0.45\\
4  & 0.64 & 0.87 & 0.81 & 0.85 & 0.45\\
8  & 0.65 & 0.89 & 0.82 & 0.86 & 0.46\\
\bottomrule
\end{tabular}
}
\end{table}

This experiment investigates the influence of the temporal encoding window on the recognition performance of the proposed RANC-based SNN architecture. The temporal encoding window, represented by the number of timesteps \(T\), determines how many temporal spike events are generated for each input sample during inference. Increasing the number of timesteps enables the network to accumulate richer temporal information and capture more discriminative spiking representations.

The experiments were conducted using the best-performing segmentation configuration identified in previous experiments, namely \((\textit{Elems}=16, \textit{Overlap}=8, \textit{Tea Out}=256)\). The timestep values were varied from \(T=1\) to \(T=8\) in order to analyze the trade-off between recognition capability and temporal computational complexity.

The experimental results in Table~\ref{tab:timestep_results} demonstrate that increasing the temporal encoding window generally improves the classification performance across all evaluation tasks. In particular, larger timestep values consistently produced higher recognition scores for \textit{Index L/R}, \textit{Thumb L/R}, \textit{Thumb/Index}, \textit{Thumb/Others}, and \textit{5 Fingers} classification tasks.

When using a single timestep (\(T=1\)), corresponding to binary spike encoding without temporal accumulation, the model achieved scores of \(0.63\) for \textit{Index L/R}, \(0.86\) for \textit{Thumb L/R}, \(0.79\) for \textit{Thumb/Index}, \(0.82\) for \textit{Thumb/Others}, and \(0.44\) for the \textit{5 Fingers} task. These results indicate that even a single-step spiking representation is capable of learning meaningful fingerprint features.

Increasing the temporal window to \(T=2\) produced small but consistent improvements across multiple tasks. The \textit{Thumb L/R} score increased from \(0.86\) to \(0.87\), while \textit{Thumb/Index} improved from \(0.79\) to \(0.80\), and \textit{Thumb/Others} improved from \(0.82\) to \(0.83\). The \textit{5 Fingers} score also increased slightly from \(0.44\) to \(0.45\). These improvements suggest that introducing additional temporal spike events allows the network to construct richer internal spiking representations.

More noticeable performance improvement was observed when increasing the timestep value to \(T=4\). Under this configuration, the \textit{Index L/R} score improved to \(0.64\), while \textit{Thumb/Index} and \textit{Thumb/Others} increased to \(0.81\) and \(0.85\), respectively. The results indicate that moderate temporal windows enable the proposed SNN architecture to better capture subtle spatial and temporal variations in capacitive fingerprint patterns through burst-based spike encoding.

The best overall performance was obtained when using \(T=8\). Under this configuration, the model achieved \(0.65\) for \textit{Index L/R}, \(0.89\) for \textit{Thumb L/R}, \(0.82\) for \textit{Thumb/Index}, \(0.86\) for \textit{Thumb/Others}, and \(0.46\) for the \textit{5 Fingers} task. Compared to the \(T=1\) configuration, the improvements demonstrate that longer temporal windows enable the network to accumulate more stable and discriminative spike-based feature representations.

The selected timestep settings were motivated by commonly adopted temporal configurations in SpikeFormer~\cite{zhou2023spikeformer} and related spiking transformer architectures, where small-to-moderate temporal windows such as \(T=1\), \(T=2\), \(T=4\), and \(T=8\) were shown to provide effective trade-offs between recognition performance and computational efficiency. Following this experimental convention enables more consistent comparison with existing spiking transformer literature while remaining suitable for lightweight neuromorphic computation.

Although increasing the timestep value generally improves performance, larger temporal windows also introduce additional computational overhead and inference latency because spike-based operations must be performed repeatedly across multiple timesteps. Moreover, the performance improvement gradually becomes smaller at higher timestep values. For example, increasing the timestep from \(T=4\) to \(T=8\) improved the \textit{Thumb/Index} score by only \(0.01\) and the \textit{5 Fingers} score by only \(0.01\). This behavior suggests diminishing representational benefits at very large temporal windows.

Overall, the experimental results demonstrate that temporal encoding plays an important role in improving the discriminative capability of the proposed SNN architecture. Moderate temporal windows such as \(T=4\) and \(T=8\) provide the most favorable balance between recognition performance and computational efficiency, while excessively large timestep values may introduce significantly higher computational cost with relatively limited additional performance improvement.

\subsection{Discussion}

The experimental results demonstrate that both spatial segmentation and temporal encoding significantly influence the performance of the proposed RANC-based SNN architecture. Among the evaluated segmentation settings, the \((16,8)\) configuration consistently achieved the best balance between recognition accuracy and model complexity, indicating that moderate segment sizes with sufficient overlap are effective for preserving local spatial continuity. Increasing the \textit{Tea Out} dimension generally improved classification performance by enabling richer spiking feature representations, although the improvement gradually saturated at higher output dimensions. In particular, increasing \textit{Tea Out} from \(128\) to \(256\) produced only marginal performance gains while substantially increasing the number of trainable parameters. 

The temporal encoding experiments further confirmed the importance of temporal spike accumulation in enhancing discriminative capability. Increasing the timestep value from \(T=1\) to \(T=8\) consistently improved recognition performance across all evaluation tasks, demonstrating the effectiveness of burst-based spike encoding for capacitive fingerprint classification. However, the improvement became progressively smaller at larger timestep values, suggesting diminishing returns beyond moderate temporal resolutions. In addition, the proposed architecture maintained relatively lightweight computational complexity while achieving competitive recognition performance through binary spike representations, discretized Tea connections, and localized parallel processing. Overall, the experimental findings validate the effectiveness of combining overlapping parallel segmentation with temporal spike encoding to achieve an efficient balance between recognition capability, representational power, and neuromorphic computational efficiency.

\section{Experimental Evaluation of Spiking Vision Transformer}

\subsection{Experimental Setup}

This section presents the experimental configuration used to evaluate the proposed Spiking Vision Transformer architecture. Similar to the previous RANC-based experiments, the evaluation focuses on both recognition performance and model complexity under different architectural and temporal configurations.

The experiments investigate the influence of three primary parameter groups adopted from the original SpikeFormer design \cite{zhou2023spikeformer}:

\begin{itemize}
    \item \textbf{Number of Spiking Transformer Encoder blocks (\(L\))}, which controls the network depth.
    
    \item \textbf{Feature embedding dimension (\(D\))}, which determines the representational capacity of the transformer features.
    
    \item \textbf{Temporal encoding window (\(T\))}, which controls the number of spike propagation timesteps.
\end{itemize}

To ensure fair comparison, only one parameter group was varied at a time while the remaining settings were fixed to their default configuration. The experiments were evaluated using the same fingerprint classification tasks and performance metrics described in the previous subsection, including \textit{Index L/R}, \textit{Thumb L/R}, \textit{Thumb/Index}, \textit{Thumb/Others}, and \textit{5 Fingers} classification accuracy.

The default configuration of the proposed Spiking Vision Transformer is summarized in Table~\ref{tab:spikformer_default_config}.

\begin{table}[h]
\centering
\caption{Default configuration of the proposed Spiking Vision Transformer}
\label{tab:spikformer_default_config}
\begin{tabular}{|c|c|}
\hline
\textbf{Parameter} & \textbf{Value} \\ \hline
Optimizer & Adam \\ \hline
Learning Rate & 0.001 \\ \hline
Maximum Epochs & 300 \\ \hline
Early Stopping Patience & 10 \\ \hline
Dropout Rate & 0.15 \\ \hline
Encoder Blocks (\(L\)) & 8 \\ \hline
Embedding Dimension (\(D\)) & 512 \\ \hline
Encoding Window (\(T\)) & 4 \\ \hline
\end{tabular}
\end{table}

\subsection{Effect of Number of Encoder Blocks}

Table~\ref{tab:spikeformer-results} presents the experimental results obtained under different Spiking Vision Transformer configurations with varying encoder depths and embedding dimensions. The results demonstrate that increasing the number of encoder blocks generally improves the recognition capability of the proposed architecture by enabling deeper hierarchical spike-based feature extraction.

Compared with the baseline model, all Spiking Vision Transformer configurations achieved noticeable performance improvements across most classification tasks. In particular, the configuration using \(8\) encoder blocks and embedding dimension \(384\) improved the \textit{Thumb L/R} accuracy from \(0.90\) to \(0.92\), while also improving the \textit{Thumb/Index} accuracy from \(0.84\) to \(0.86\). These results indicate that the transformer-based spiking representation is capable of learning more discriminative spatial relationships than the baseline architecture.

When the embedding dimension was fixed at \(D=512\), increasing the number of encoder blocks from \(L=6\) to \(L=8\) consistently improved performance across all evaluation tasks. Specifically, \textit{Index L/R} accuracy increased from \(0.67\) to \(0.70\), while \textit{Thumb/Others} accuracy improved from \(0.87\) to \(0.90\). This behavior suggests that deeper transformer structures allow the network to capture richer contextual dependencies between fingerprint regions through multiple stages of spike-based self-attention.

Further increasing the encoder depth to \(L=10\) produced additional improvement in several tasks. Under the \((L=10, D=512)\) configuration, the model achieved the highest \textit{Thumb/Others} accuracy of \(0.92\) and the best \textit{5 Fingers} classification accuracy of \(0.48\). These results demonstrate that deeper spiking transformer architectures are particularly beneficial for more challenging multi-class classification scenarios requiring stronger global feature aggregation.

The results also indicate that the improvement gradually saturates at higher model capacities. Although the \(L=10\) configuration outperformed shallower models in some tasks, the performance gain compared with \(L=8\) remained relatively small. This suggests that excessively deep transformer architectures may introduce additional computational complexity while providing only marginal improvement in recognition capability.

Overall, the experiments confirm that encoder depth plays an important role in the representational power of the proposed Spiking Vision Transformer. Moderate-to-deep configurations, particularly those using \(8\) to \(10\) encoder blocks, provide the most favorable balance between classification performance and architectural complexity for capacitive fingerprint recognition.

\begin{table*}[t]
\centering
\caption{Performance of Spiking Vision Transformer under different encoder depth and embedding dimension configurations (\(T=4\)).}
\resizebox{1\textwidth}{!}{
\begin{tabular}{ccccccc}
\toprule
\multicolumn{2}{c}{\textbf{Architecture}} & \textbf{Index L/R} & \textbf{Thumb L/R} & \textbf{Thumb/Index} & \textbf{Thumb/Others} & \textbf{5 Fingers}\\
\cmidrule(lr){1-2}
Encoder Blocks (\(L\)) & Embedding Dimension (\(D\)) & & & & &\\
\midrule
\multicolumn{2}{c}{Baseline Model} & 0.65 & 0.90 & 0.84 & 0.86 & 0.46 \\
8 & 384 & 0.65 & 0.92 & 0.86 & 0.86 & 0.45 \\
6 & 512 & 0.67 & 0.92 & 0.87 & 0.87 & 0.45 \\
8 & 512 & 0.70 & 0.94 & 0.89 & 0.90 & 0.46 \\
10 & 512 & 0.70 & 0.94 & 0.89 & \textbf{0.92} & \textbf{0.48} \\
8 & 768 & \textbf{0.71} & \textbf{0.95} & \textbf{0.90} & 0.92 & 0.47 \\
\bottomrule  
\end{tabular}
}
\label{tab:spikeformer-results}
\end{table*}

\subsection{Effect of Changing Embedding Dimensions}

The experimental results further demonstrate that the embedding dimension significantly affects the representational capability of the proposed Spiking Vision Transformer. Increasing the embedding dimension generally improved classification performance across most evaluation tasks by allowing the network to encode richer spike-based feature representations.

When the number of encoder blocks was fixed at \(L=8\), increasing the embedding dimension from \(D=384\) to \(D=512\) consistently improved recognition accuracy. In particular, \textit{Index L/R} accuracy increased from \(0.65\) to \(0.70\), while \textit{Thumb/Others} accuracy improved from \(0.86\) to \(0.90\). Similarly, the \textit{Thumb/Index} classification accuracy increased from \(0.86\) to \(0.89\). These improvements indicate that larger embedding spaces enable the transformer to preserve more discriminative spatial and temporal spike features.

Further increasing the embedding dimension to \(D=768\) produced the best overall performance in several classification tasks. Under the \((L=8, D=768)\) configuration, the model achieved the highest \textit{Index L/R} accuracy of \(0.71\), the highest \textit{Thumb L/R} accuracy of \(0.95\), and the best \textit{Thumb/Index} accuracy of \(0.90\). These results suggest that larger embedding dimensions improve the ability of the self-attention mechanism to model complex relationships between fingerprint regions.

However, the experimental results also indicate diminishing performance gains at very large embedding dimensions. Although the \(D=768\) configuration achieved slightly better performance than \(D=512\), the improvement remained relatively modest in several tasks. For example, the \textit{Thumb/Others} accuracy remained unchanged at \(0.92\), while the \textit{5 Fingers} accuracy improved only marginally from \(0.46\) to \(0.47\). This suggests that increasing embedding dimensionality beyond moderate values may introduce additional computational overhead with limited practical benefit.

Compared with the baseline model, all transformer-based configurations achieved substantially higher performance, demonstrating the effectiveness of spike-based transformer representations for capacitive fingerprint recognition. The results confirm that larger embedding dimensions improve feature diversity and contextual representation learning within the spiking self-attention framework.

Overall, the experiments show that embedding dimension is a critical factor influencing the representational capacity of the Spiking Vision Transformer. Moderate-to-large embedding dimensions, particularly \(D=512\) and \(D=768\), provide the most favorable balance between recognition performance and architectural efficiency.

\subsection{Effect of Temporal Encoding Window}

\begin{table*}[t]
\centering
\caption{Performance of the Spiking Vision Transformer under Different Temporal Encoding Windows.}
\resizebox{1\textwidth}{!}{
\begin{tabular}{cccccc}
\toprule
Time Step & \textbf{Index L/R} & \textbf{Thumb L/R} & \textbf{Thumb/Index} & \textbf{Thumb/Others} & \textbf{5 Fingers}\\
\midrule
Baseline Model & 0.65 & 0.90 & 0.84 & 0.86 & 0.46 \\
1 & 0.65 & 0.89 & 0.86 & 0.86 & 0.45 \\
2 & 0.67 & 0.92 & 0.87 & 0.86 & 0.46 \\
4 & 0.70 & 0.95 & 0.89 & 0.89 & 0.48 \\
8 & \textbf{0.71} & \textbf{0.95} & \textbf{0.90} & \textbf{0.90} & \textbf{0.48} \\
\bottomrule  
\end{tabular}
}
\label{tab:spikeformer-timestep-results}
\end{table*}

This experiment investigates the influence of temporal encoding on the proposed Spiking Vision Transformer. To ensure fair comparison across different temporal resolutions, all timestep experiments were conducted using the same transformer configuration with \(L=8\) encoder blocks and embedding dimension \(D=512\), while only the number of encoding timesteps was varied.

The experimental results demonstrate that increasing the temporal encoding window generally improves the recognition capability of the proposed architecture. Compared with the baseline model, all spiking transformer configurations achieved competitive or superior performance across most evaluation tasks, indicating that temporal spike-based processing effectively enhances feature representation learning.

When using a single timestep (\(T=1\)), the model achieved \(0.89\) accuracy for the \textit{Thumb L/R} task and \(0.86\) for \textit{Thumb/Index}. Although the performance remained comparable to the baseline model, the limited temporal resolution restricted the ability of the network to fully exploit temporal spike dynamics.

Increasing the timestep to \(T=2\) produced consistent improvements across several tasks. In particular, \textit{Index L/R} accuracy improved from \(0.65\) to \(0.67\), while \textit{Thumb L/R} increased from \(0.89\) to \(0.92\). These results suggest that introducing additional temporal information enables the transformer to learn richer spike-based contextual representations.

More significant improvement was observed when increasing the temporal window to \(T=4\). Under this configuration, the model achieved \(0.70\) accuracy for \textit{Index L/R}, \(0.95\) for \textit{Thumb L/R}, and \(0.89\) for both \textit{Thumb/Index} and \textit{Thumb/Others}. The \textit{5 Fingers} classification accuracy also increased to \(0.48\), outperforming the baseline model. These results indicate that moderate temporal windows allow the spiking transformer to better capture temporal correlations and subtle fingerprint variations.

Further increasing the temporal window to \(T=8\) produced only marginal improvement. Although the model achieved the best overall performance, including \(0.71\) for \textit{Index L/R} and \(0.90\) for both \textit{Thumb/Index} and \textit{Thumb/Others}, the improvement over \(T=4\) remained relatively small. This behavior suggests diminishing returns at higher timestep values, where additional temporal information contributes limited representational benefit.

Overall, the experiments confirm that temporal encoding plays a crucial role in the performance of the proposed Spiking Vision Transformer. Moderate timestep values such as \(T=4\) and \(T=8\) provide the most favorable balance between recognition accuracy and computational efficiency, while excessively small temporal windows limit the ability of the network to exploit spike-based temporal dynamics effectively.

\subsection{Discussion}

The experimental results demonstrate that the proposed Spiking Vision Transformer consistently outperforms the baseline model across most fingerprint classification tasks. The integration of spike-based self-attention enables the network to capture richer spatial and temporal relationships between fingerprint regions. Increasing the number of encoder blocks generally improves recognition performance, particularly for more complex tasks such as \textit{Thumb/Others} and \textit{5 Fingers}. However, the improvement gradually saturates at deeper configurations, indicating diminishing returns in performance compared with the additional computational cost.

Similarly, larger embedding dimensions improve the representational capability of the transformer by allowing more discriminative spike-based feature encoding. The best overall performance was achieved using moderate-to-large embedding dimensions such as \(D=512\) and \(D=768\). The temporal encoding experiments also confirm the importance of spike-based temporal dynamics. Increasing the timestep from \(T=1\) to \(T=4\) significantly improved classification accuracy, while further increasing to \(T=8\) produced only marginal gains. Overall, the results indicate that moderate encoder depth, embedding dimension, and temporal window size provide the best balance between recognition performance and computational efficiency for capacitive fingerprint recognition.

\end{document}